\LARGE \bf IMU-based pose-estimation for spherical robots with limited resources

### 2021-05-31 mfi21/ieeeconf/root.tex
changed photo
now here comes the actual input for 3d mapping

This paper introduces an IMU-based orientation and position algorithm for spherical robots with limited computational power and the need for real-time estimations.
Also, the algorithm does not rely on one specific locomotion approach, but it assumes a rotation of the IMU together with at least the shell of the robot, if not the whole robot. 
Such as the prototype developed at the computer science chair VII of the University of Wuerzburg,  depicted in Figure \ref{fig:prototype}, which is inspired by previous work such as Daedalus \cite{rossi2021daedalus}.
The evaluations made in this paper were done with the setup of the IMUs of this prototype.
The Livox Mid-100 was used as laser scanner.
Due to the lack of a locomotion system, the sphere needs to be rolled externally. 
The evaluation will also show and discuss the resulting point clouds of this prototype with the presented pose estimation.
\begin{figure}[b]
\includegraphics[width=0.484\linewidth]{./graphics/photoReal1.jpeg}
\includegraphics[width=0.50\linewidth]{./graphics/photoReal2.jpeg}
\caption{Prototype to test the algorithm in an envisaged technical environment. The main payload is the Livox Mid-100 laser-scanner. For pose-estimation, three IMUs of the manufacturer Phidget are placed on the middle plate and a Raspberry Pi 4 for the calculations. On the top are two batteries, and on the bottom one voltage stabilizer and breakout box of the laser-scanner. }
\label{fig:prototype}
\end{figure}

@@

%TODO Fabian, put input of laserscanns here

>>>>

This paper introduces an IMU-based orientation and position algorithm for spherical robots with limited computational power and the need for real-time estimations.
Also, the algorithm does not rely on one specific locomotion approach, but it assumes a rotation of the IMU together with at least the shell of the robot, if not the whole robot. 
Such as the prototype developed at the computer science chair VII of the University of Wuerzburg,  depicted in Figure \ref{fig:prototype}, which is inspired by previous work such as Daedalus \cite{rossi2021daedalus}.
The evaluations made in this paper were done with the setup of the IMUs of this prototype.
The Livox Mid-100 was used as laser scanner.
Due to the lack of a locomotion system, the sphere needs to be rolled externally. 
The evaluation will also show and discuss the resulting point clouds of this prototype with the presented pose estimation.
\begin{figure}[b]
\includegraphics[width=0.49\linewidth]{./graphics/photoReal1.jpeg}
\includegraphics[width=0.50\linewidth]{./graphics/photoReal3.jpeg}
\caption{Prototype to test the algorithm in an envisaged technical environment. The main payload is the Livox Mid-100 laser-scanner. For pose-estimation, three IMUs of the manufacturer Phidget are placed on the middle plate and a Raspberry Pi 4 for the calculations. On the top are two batteries, and on the bottom one voltage stabilizer and breakout box of the laser-scanner. }
\label{fig:prototype}
\end{figure}

@@

\subsection{Impact on 3D mapping}

\begin{figure}
\begin{center}
\includegraphics[width=\linewidth]{./graphics/jasperhome1.png} \\
\vspace{0.2cm}
\includegraphics[width=\linewidth]{./graphics/jasperhome1madw.png}
\caption{(Top) Resulting 3D map using the presented algorithm for pose filtering. (Bottom) Resulting 3D map using Madgwick + double integrator for pose filtering. Both images were shot from the exact same point of view. }
\label{fig:mapping}
\end{center}
\end{figure}

One major design driver for the presented algorithm is the suitability for spherical SLAM purposes.
Therefore, a pose estimate is prefered to be slightly wrong, if it is more continuous, i.e. less noisy and with less jumping values, than the exact pose would be.

To test the impact of the filter on 3D mapping, we used the prototype in Figure \ref{fig:prototype}.
After 2 s of initial standstill, the sphere rolled aprox. 4 m within 5 s in an indoor environment from one corner of the room to the opposing one.
We extracted the laserscan data, as well as the filtered pose using the presented algorithm.
Without further pre- or post-processing, the pose data is combined with the laserscan data to create a 3D map.
Figure \ref{fig:mapping} compares the result with another 3D map, which is generated the exact same way, but using a different pose filter.
Namely a combination of the Madgwick filter for orientation and a double integration of the accelerometer mean for position.
The result of this filter clearly got improved by the presented algorithm as it provides a better pose estimation.
The more detailed part of the map, as seen in the bottom left corner of the top image of Figure \ref{fig:mapping}, corresponds to the initial standstill.
It is more detailed than the rest of the map because of the non-repeating flower-shaped scanning pattern, where point density increases with time.
Note that the same detailed part is also present in the bottom image.
However, it is out of view (at the bottom right corner) as the Madgwick filter misses the initial pose estimation by a lot. 
The presented algorithm also performs better when the sphere is moving, as the walls and the ceiling are more recognizable, which is favourable for many SLAM algorithms.
